---
title: "Assignment-3"
author: "Harika"
date: "2023-11-05"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r}
library(caret)
library(ggplot2)
library(ISLR)
library(class)
library(reshape2)
#install.packages("melt")
library(melt)
```

1.DATA LOADING

```{r}
File_path <- "C:/Users/Harika/Downloads/UniversalBank-1.csv"
```

```{r}
###specified the file path above that to be loaded.
```

```{r}
UniversalBank_1 <- read.csv("C:/Users/Harika/Downloads/UniversalBank-1.csv")
```

```{r}
### Loaded the dataset above
```

```{r}
UniversalBank_1$Personal.Loan <- as.factor(UniversalBank_1$Personal.Loan)
UniversalBank_1$Online <- as.factor(UniversalBank_1$Online)
UniversalBank_1$CreditCard <- as.factor(UniversalBank_1$CreditCard)
```

```{r}
summary(UniversalBank_1)
```
```{r}
###the above data represents summary for the given dataset.
```


2.DATA PARTITIONING (60:40)

```{r}
set.seed(23)
```

```{r}
###to set a seed for reproducibility.
```

```{r}
SplitData <- createDataPartition(UniversalBank_1$Personal.Loan, p=0.60, list = FALSE)
train_data <- UniversalBank_1[SplitData,]
test_data <- UniversalBank_1[-SplitData,]
```

```{r}
###the data is now splitted into training (60%) and testing (40%) sets above.
```

```{r}
dim(train_data)
dim(test_data)
```
```{r}
###to check the dimensions of the training and testing sets above.
```

>A. TO CREATE A PIVOT TABLE

```{r}
Pivot_Table1 <- ftable(train_data$CreditCard, train_data$Personal.Loan, train_data$Online)
Pivot_Table1
```
```{r}
###the pivot table created above.
```

>B. Based on the pivot table created, we can determine the probability that this customer would accept the loan offer, when equals 53/(53+497) ~ 0.096.

>C. Two separate pivot tables were created using the training data. Where one will have the internet (columns) are a function of the personal loan (rows), and whereas the credit card is a function of the other.

```{r}
Melt_1 <- melt(train_data,id=c("Personal.Loan"),variable="Online")
Melt_2 <- melt(train_data,id=c("Personal.Loan"),variable="CreditCard")
```
```{r}
cast1 = dcast(Melt_1, Personal.Loan~Online)
```
```{r}
cast2 = dcast(Melt_2, Personal.Loan~CreditCard)
```
>D. Calculating specified amounts P(A/B) signifies the probability which A will occur given B.

```{r}
ftable(train_data$Personal.Loan, train_data$Online)
```
```{r}
ftable(train_data$Personal.Loan, train_data$CreditCard)
```
>ftable(train_data[,10]) #10 is Personal.Loan column.

1. P(CC = 1 | Loan = 1) = (92/92+196) = 0.319
2. P(Online = 1 | Loan = 1) = (167/167+121) = 0.579
3. P(Loan = 1) = (288/288+2712) = 0.096
4. P(CC = 1 | Loan = 0) = (812/812+1900) = 0.299
5. P(Online = 1 | Loan = 0) = (1624/ 1624+1088) = 0.598
6. P(Loan = 0) = (2712/ 2712+288) = 0.904

>E. Using quantities computed above to compute the naive bayes probability P(Loan=1|CC=1, Online=1).

(0.319*0.579*0.096)/(0.319*0.579*0.096)+(0.299*0.598*0.904) ~ 0.098

>F.In question B, we calculated a probability value of 0.096, and in the previous question, we found a probability value of 0.098. These values have a slight difference only. As of question E, we took into account more dependent information, but it seems that the value from question B is more accurate and specific comparitively.

>G.Implementing Naives Bayes below

```{r}
#install.packages("naivebayes")
library(naivebayes)
```

```{r}
Naive_Bayes_Model <- naive_bayes(Personal.Loan~Online+CreditCard,data = train_data)
Naive_Bayes_Model
```
```{r}
###
```


